---
title: "Spotify Wrapped (Steph's Version)"
output: html_document
author: Kathleen Dang
date: "2022-12-22"
---
## Data preparation
#### Import libraries
```{r, results= 'hide', warning=FALSE, message=FALSE}
library(tidyverse)
library(lubridate)
library(jsonlite)
library(scales)
library(ggrepel)
```

#### Load in and merge data files
```{r}
# replace with specific data directory
setwd("C:/Users/kat4538/Documents/spotify/steph's data")

# can add or remove n files depending on number you have
# e.g. c("StreamingHistory0.json","StreamingHistory1.json", ... ,"StreamingHistoryn.json")
stream <- c("StreamingHistory0.json","StreamingHistory1.json", "StreamingHistory2.json")

streamHistory = do.call("rbind", lapply(stream, function(f) fromJSON(f)))
```

#### Create tibble with additional time variables (derived from the original variables provided)
```{r}
spotify = streamHistory %>% 
  as_tibble() %>% 
  mutate_at("endTime", ymd_hm) %>% 
  mutate(endTime = endTime - hours(6)) %>% 
  mutate(date = floor_date(endTime, "day") %>% 
  as_date, seconds = msPlayed / 1000, minutes = seconds / 60, hours = minutes / 60) %>%
  mutate(month = format(date, "%B")) 

# check first row
spotify %>% head(n=1)
```
## Top track of every month
#### What song did you listen to most each month?
```{r, fig.align='center'}
# create vector of month names
order = c('January', 'February', 'March', 'April', 'May', 'June', 'July', 'August', 'September', 'October', 'November', 'December')

months = spotify %>% filter(date >= "2022-01-01") %>%
  group_by(trackName, month) %>%
  summarize(totalMin = sum(minutes), .groups = 'drop') %>%
  group_by(month) %>%
  slice_max(totalMin, n = 1) %>%
  mutate(month = factor(month, levels = order)) %>%
  arrange(month)
knitr::kable(months)

options(repr.plot.res = 300, repr.plot.height = 5, repr.plot.width = 7)
ggplot(months, aes(x = month, y = totalMin)) + geom_point(color = "seagreen3") +
 geom_text_repel(data = months, size= 3, aes(label= trackName), 
                        fontface = 'bold', force= 1, point.padding = unit(4, 'lines')) +
  scale_y_continuous(limits=c(0, 200), n.breaks = 10) + 
 labs(x= "Month", y= "Minutes of music playback", 
             title = "Most played track each month\n(January-November 2022)") +
  theme_grey() + theme(axis.text.x = element_text(angle = 45,  hjust= 1), 
              plot.title = element_text(hjust = 0.5))
```



## Streaming activity by week over the months
#### What's the trend of streaming activity over time?
```{r, fig.align='center'}
streamingHours = spotify %>% filter(date >= "2022-01-01") %>%
  group_by(date) %>% 
  group_by(date = floor_date(date, "week")) %>%
  summarize(hours = sum(minutes) / 60, minutes= sum(seconds)/60) %>% 
  arrange(date)

ggplot(streamingHours, aes(x = date, y = hours, color= minutes)) + geom_line() +
  labs(x= "Month", y= "Hours of music playback", 
       title = "Playback activity per week over the months \n(January-November 2022)", color= "Minutes") +
  scale_color_continuous(labels = comma) + scale_x_date(date_labels = "%b", date_breaks = "1 month") +
  theme_minimal() + theme(plot.title = element_text(hjust = 0.5))
```


## Heatmap of streaming activity by time of day and day of week
#### When do you tend to listen to music?
```{r, fig.align='center'}
hoursDay = spotify %>% 
  group_by(date, hour = hour(endTime), weekday = wday(date, label = TRUE))%>% 
  summarize(minutesListened = sum(minutes))

hoursDay_df = hoursDay %>% 
  group_by(weekday, hour) %>% 
  summarize(minutes = sum(minutesListened)) 

ggplot(hoursDay_df, aes(x = hour, weekday, fill= minutes)) + geom_tile() +
  scale_fill_gradient(low = "yellow", high = "red") + 
  labs(x= "Time of the day (Hour 0-24)", y= "Weekday", 
       title= "What day of the week and time of day \ndo I listen to music the most on Spotify?", fill = "Minutes") +
  scale_x_continuous(breaks = c(0,4,8,12,16,20,24)) + theme_classic() + 
  theme(plot.title = element_text(hjust = 0.5))
```


## Top artists of the year
#### Which artists did you listen to most this year?
```{r, fig.align='center'}
artist = spotify %>% 
  group_by(artistName) %>% 
  summarize(minutesListened = sum(minutes), hoursListened = sum(hours)) %>% 
  slice_max(order_by = minutesListened, n=10)

ggplot(artist, aes(x = reorder(artistName, desc(minutesListened)), y = minutesListened)) + geom_col(fill= "hotpink1") +
  labs(x= "Artist", y= "Minutes of music playback", 
             title = "My top 10 most streamed artists on Spotify") + 
  scale_y_continuous(labels = comma) + theme_bw() + theme(axis.text.x = element_text(angle = 45,  hjust= 1))

```


## Top songs by one artist
#### Which songs did you listen to most by a specific artist?
```{r, fig.align='center'}
# filter by specific artist name

# Phil Collins
spotify %>% filter(artistName == "Phil Collins") %>% summarize(totalPhil = sum(minutes))

phil = spotify %>%
  filter(artistName == "Phil Collins") %>%
  group_by(trackName) %>%
  summarize(minutes = sum(minutes), .groups = 'drop') %>%
  arrange(minutes) %>%
  slice_max(minutes, n = 10) %>%
  mutate(trackName = fct_inorder(trackName))

ggplot(phil, aes(x=trackName, y= minutes)) + geom_col(fill='darkolivegreen3') + 
  theme_light() + theme(axis.text.x = element_text(hjust= 0.5)) +
  scale_x_discrete(labels = label_wrap(10)) +
  scale_y_continuous(labels = comma) + labs(x= "Track name", y= "Minutes of music playback", 
             title = "My top 10 most listened tracks by Phil Collins")

# Brooks and Dunn
spotify %>% filter(artistName == "Brooks & Dunn") %>% summarize(totalPhil = sum(minutes))

bd = spotify %>%
  filter(artistName == "Brooks & Dunn") %>%
  group_by(trackName) %>%
  summarize(minutes = sum(minutes), .groups = 'drop') %>%
  arrange(minutes) %>%
  slice_max(minutes, n = 10) %>%
  mutate(trackName = fct_inorder(trackName))

ggplot(bd, aes(x=trackName, y= minutes)) + geom_col(fill='darkolivegreen3') + 
  theme_light() + theme(axis.text.x = element_text(angle = 45,  hjust= 1)) +
  scale_y_continuous(labels = comma) + labs(x= "Track name", y= "Minutes of music playback", 
                                            title = "My top 10 most listened tracks by Brooks & Dunn")
```



## Total minutes spent streaming music
#### How many minutes did you listen to music this year?
```{r}
totalMin <- spotify %>% filter(date >= "2022-01-01") %>% 
            summarize(totalMinutes=round(sum(minutes),0), percentage=totalMinutes/525600,
                      percent = round(percentage*100, 0))
paste0("You listened to ", totalMin$totalMinutes, " minutes of music. That's ",
        totalMin$percent, "% of your year spent bumping to some bangers!")

```

